\documentclass[10pt]{article}
\usepackage[margin=2cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{fancyvrb}

\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue}

\pagestyle{empty}

\begin{document}

\begin{center}
{\Large\bfseries Substrate Scanning Engine --- Quick Start}\\[4pt]
{\small v3.0.0 \quad Ricardo Hern\'andez Reveles \quad ORCID: 0009-0001-1993-1471}
\end{center}

\vspace{6pt}
\hrule
\vspace{8pt}

\noindent\textbf{What it does.}\quad
Takes any system with a linear operator and discrete spectrum, extracts
31 deterministic geometric features (29 core + 2 Euler-specific) per query point,
plus 10 CF-integer features for diophantine analysis of zeros. No statistical
assumptions. NumPy, SciPy, mpmath.

\vspace{4pt}
\noindent\textbf{The fundamental equation:}\quad
$p^{-s} = p^{-\sigma}\cdot e^{-i\ln(p)\cdot t}$

\vspace{6pt}
\noindent\textbf{1. Install}
\begin{Verbatim}[fontsize=\small, frame=single]
git clone https://github.com/rhrev/substrate-scanning-engine.git
cd substrate-scanning-engine
pip install -r requirements.txt   # numpy>=1.24, scipy>=1.10, mpmath>=1.3
\end{Verbatim}

\vspace{4pt}
\noindent\textbf{2. Run the core engine (Riemann zeta)}
\begin{Verbatim}[fontsize=\small, frame=single]
python pipelines/geometric_engine.py
\end{Verbatim}
This prints: 31-feature vector (29 core + 2 Euler-specific), three-scale decomposition, envelope
factorization, potential landscape, Weil spectrum, and channel orthogonality
for $\zeta(s)$, $L(s,\chi)$, and $L(s,E)$.  New in v3.0.0: run
\verb|python pipelines/cf_features.py| and \verb|python pipelines/zero_locator.py|
for the CF-integer and zero-localization pipelines.

\vspace{4pt}
\noindent\textbf{3. Run any application engine}
\begin{Verbatim}[fontsize=\small, frame=single]
python applications/sparam_engine.py       # S-parameters (1709/s)
python applications/motor_diagnosis.py     # Vibration faults (1132/s)
python applications/phonon_engine.py       # Materials screening (478/s)
python applications/crypto_curve_engine.py # Elliptic curve audit (274/s)
\end{Verbatim}
Each engine generates synthetic data, extracts 29 features, classifies,
and prints a cross-domain mapping to~$\zeta$.

\vspace{4pt}
\noindent\textbf{4. Use as a library}
\begin{Verbatim}[fontsize=\small, frame=single]
import sys; sys.path.insert(0, 'pipelines')
from geometric_engine import GeometricEngine, RiemannZeta

ep = RiemannZeta()
engine = GeometricEngine(ep)
vec = engine.feature_vector(0.5, 14.134725, 30)  # 31 floats
print(f"Features: {len(vec)}, phase coherence: {vec[23]:.4f}")
\end{Verbatim}

\vspace{4pt}
\noindent\textbf{5. Locate zeros (Riemann zeta)}
\begin{Verbatim}[fontsize=\small, frame=single]
import sys; sys.path.insert(0, 'pipelines')
from zero_locator import ZeroLocator
from geometric_engine import RiemannZeta
from mpmath import mp; mp.dps = 50

loc = ZeroLocator(RiemannZeta())
zeros = loc.scan(10, 50, dps=50, cf_depth=25)
for z in zeros:
    print(f"gamma={float(z.gamma):.10f}, residual={z.residual:.2e}")
\end{Verbatim}
For Dirichlet/EC zeros, use \verb|loc.from_values([6.02, 10.24], dps=50)|
with externally computed zeros (e.g., from LMFDB).

\vspace{4pt}
\noindent\textbf{6. Extract CF-integer features}
\begin{Verbatim}[fontsize=\small, frame=single]
from cf_features import CFFeatureExtractor, batch_features
from mpmath import mp, zetazero; mp.dps = 100

gammas = [zetazero(n).imag for n in range(1, 21)]
per_f, cross_f, terms = batch_features(gammas, K=50, dps=100)
# per_f: (20, 10) — 10 features per zero
# cross_f: (190, 4) — 4 pairwise features
\end{Verbatim}

\vspace{4pt}
\noindent\textbf{7. Explore the visualizations}\\
Open any HTML file in \texttt{visualizations/} in a browser. Requires
internet for Three.js r128 CDN. 17 interactive 3D scenes: toroidal
embeddings, potential landscapes, operator spectra, geodesics.

\vspace{4pt}
\noindent\textbf{8. Add a new domain}\\
See \texttt{CONTRIBUTING.md}. Short version: write a standalone
\texttt{.py} in \texttt{applications/} that extracts 29 features from
your spectrum, include a synthetic demo and cross-domain mapping.
Reference template: \texttt{sparam\_engine.py}.

\vspace{8pt}
\hrule
\vspace{6pt}

\noindent\textbf{All 12 application engines:}
sparam (SI), motor\_diagnosis (vibration), graph (networks),
ast (code), cavity (resonators), phonon (materials),
colorimetry (spectra), wavelet\_3d (volumes), turbulence (cascades),
image\_spectral (forensics), pathfinding (navigation),
crypto\_curve (ECC audit).

\vspace{4pt}
\noindent\textbf{Read \S0 of the paper first.}
Seven overclaims were produced and corrected. The errata-first protocol is
the methodology's signature.

\vspace{4pt}
\noindent\emph{Democratising scepticism is worth more than democratising computation.}

\vspace{4pt}
\noindent
License: AGPL-3.0-or-later (code), CC-BY-4.0 (papers).
DOI: \href{https://doi.org/10.5281/zenodo.19104086}{10.5281/zenodo.19104086}

\end{document}
